Dark energy is based on a math error from 1930

### 2024-12-28 paper/paper.tex

\begin{abstract}
  We show that the magnitude of distant objects has been calculated incorrectly
  since 1930. We describe how this math error has propagated for over nine
  decades. Finally, we explore the consequences of correcting this flaw and
  resolve two of the preeminent mysteries in cosmology: dark energy is not
  supported by observations of Type Ia supernovae, and the Hubble tension is
  due to a calculation error.
\end{abstract}

% keywords can be removed
\keywords{Cosmological Parameters \and K-corrections \and Dark Energy \and Luminosity Distance \and Hubble tension}

@@

The measurement of Type Ia supernovae is one of the primary sources of data for
cosmological models. The measurements involve estimating the peak magnitude in
the B filter \citet{riess1998} for the rest frame, meaning we want to know how
bright the SnIa would be if its light didn't undergo redshift. Redshift causes
the light that would be measured by the B filter in a rest frame to be shifted
to another filter. In order to observe a SnIa in an arbitrary filter and report
the magnitude in the B filter we use a technique called K-corrections.

The first formal derivation of K-corrections was performed by
\citet{tolman1930}, and it failed to account for bandwidth stretching, one of
the dimming effects of redshift. This error was noted by \citet{desitter1934},
but in \citet{hubble1935} the discrepancy was noted and ignored.
\citet{oke1968} rederived K-corrections, but they failed to account for time
dilation, one of the other dimming effects of redshift. In \citet{kim1996},
K-corrections were extended to handle additional cross-filter comparisons and
address zero-point corrections, but they started their derivation by
referencing the incorrect equation from \citet{oke1968}. \citet{hsiao2007}
expanded on the work done by \citet{kim1996} by providing expected SnIa
template spectra. Finally, \citet{burns2010} implemented the popular software
package Sn(OO)py which implemented the SnIa template spectra into
K-corrections. This software package computes K-corrections differently from
prior work, and in order to match prior values, the code adds an extra term.
This history is explored in more depth in Section \ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect compounds for greater
distances. Up until \citet{riess1998}, we didn't have distant enough
observations for this error to matter much for cosmology models, but in the
late 1990s, it was clear that our measurements for distance and redshift were
not linear. This observation led to a model that utilized a cosmological
constant and dark energy in order to explain the non-linear distance-redshift
graph. We will show in Section \ref{sec:correction} that correcting the flaw
with K-corrections leads to a linear graph that does not need to rely on a
cosmological constant.

Since \citet{planck2020}, it has been evident that something was missing with
our understanding of cosmology. This research used an alternative technique
based on measurements of the cosmic microwave background to measure the Hubble
constant, but as summarized in \citet{perivolaropoulos2022}, this measurement
was incompatible with the Hubble constant measured from SnIa. As we will
discuss in Section \ref{sec:tension}, fixing the error with K-corrections
produces a measurement of the Hubble constant that is compatible with models
based on the cosmic microwave background.

@@

\begin{equation}
 F'\left(\frac{\lambda}{1+z}\right) = \frac{\left(\frac{\lambda}{1 + z}\right) F\left(\frac{\lambda}{1+z}\right)}{hc}.
\end{equation}

@@

\begin{equation}
\begin{aligned}
\label{eq:redshifted_density}
  R'(\lambda) &= F'(\lambda / (1+z)) \times \frac{1}{(1 + z)^2} \\
  R(\lambda) &= \frac{F(\lambda / (1+z))}{(1+z)hc} \times \frac{1}{(1 + z)^2} .
\end{aligned}
\end{equation}

@@

\begin{equation}
\begin{aligned}
  K'_{xy} &= \frac{\mathcal{F}_x P'_x}{\mathcal{F}_y P'_y} \\
         &= \frac{P'_x \int F'(\lambda) S_x d\lambda}
                 {P'_y \int R'(\lambda) S_y d\lambda} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x d\lambda}
                   {\int F'(\lambda / (1+z)) \times \frac{1}{(1 + z)^2} d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1+z)^2 \times
              \frac{\int \frac{\lambda F(\lambda)}{hc} S_x d\lambda}
                   {\int \frac{F(\lambda / (1+z))}{(1+z)hc} S_y d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1 + z)^2 \times
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} .
\end{aligned}
\end{equation}

@@

\begin{equation}
\begin{aligned}
\label{eq:K2k}
  K_{xy} &= 2.5\text{log}(K'_{xy}) \\
         &= 2.5\text{log}\left(
            \frac{P'_x}{P'_y} \times (1 + z)^2 \times
            \frac{\int \lambda F(\lambda) S_x d\lambda}
                 {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda}\right) \\
         &= 2.5 \left(
            \text{log} \left( \frac{P'_x}{P'_y} \right)
            + \text{log}( {(1 + z)^2})
            + \text{log}\left( \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda}
            \right) \right) \\
         &= 2.5 \text{log} \left( \frac{P'_x}{P'_y} \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} \right) \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} \right)
            - P_x + P_y .
\end{aligned}
\end{equation}

@@

\begin{equation}
\begin{aligned}
  K_{xy} &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} \right)
            - P_x + P_y \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} \right)
            + 2.5 \text{log} \left( \int \frac{\lambda}{hc} \mathcal{Z}(\lambda) S_x(\lambda) d\lambda \right)
            - 2.5 \text{log} \left( \int \frac{\lambda}{hc} \mathcal{Z}(\lambda) S_y(\lambda) d\lambda \right) \\
         &= 2.5 \text{log} \left(
              \frac{\int \lambda \mathcal{Z}(\lambda) S_x(\lambda) d\lambda}
                   {\int \lambda \mathcal{Z}(\lambda) S_y(\lambda) d\lambda}
             \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x d\lambda}
                   {\int \frac{\lambda}{1+z} F(\lambda / (1+z)) S_y d\lambda} \right) .
\end{aligned}
\end{equation}

@@

\begin{itemize}
  \item The sign is different for the first term. This will have a subtle but
  frustrating effect for real values. Typically the zero-point values are
  similar, so this term is close to 0. However, the term isn't always zero. As
  we observe more distant supernovae, it is useful to switch our observational
  filter to a filter that is sensitive to longer wavelengths. Each time we make
  this switch we introduce a small but consistent bias.

  This effect by itself can alter estimates of $H_0$ by several km/s / Mpc.
  While we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we aren't terribly confident in the numbers because we
  were unable to correct for the zero-point errors.

>>>>

\begin{abstract}
  We show that the magnitude of distant objects has been calculated incorrectly
  since 1930. We describe how this math error in the formula for K-corrections
  has propagated for over nine decades before re-deriving the K-corrections
  equation. Finally, we explore the consequences of correcting this flaw and
  resolve two of the preeminent mysteries in cosmology: dark energy is not
  supported by observations of Type Ia supernovae, and the Hubble tension is
  due to a calculation error.
\end{abstract}

@@

The measurement of Type Ia supernovae is one of the primary sources of data for
cosmological models. The measurements involve estimating the peak magnitude in
the B filter \citet{riess1998} for the rest frame, meaning we want to know how
bright the SnIa would be if its light did not undergo redshift. Redshift causes
the light that would be measured by the B filter in a rest frame to be shifted
to another filter. In order to observe a SnIa in an arbitrary filter and report
the magnitude in the B filter we use a technique called K-corrections.

The first formal derivation of K-corrections was performed by
\citet{tolman1930} and it failed to account for bandwidth stretching, one of
the dimming effects of redshift. This error was noted by \citet{desitter1934},
but in \citet{hubble1935} the discrepancy was noted and ignored.
\citet{oke1968} rederived K-corrections, but they failed to account for time
dilation, one of the other dimming effects of redshift. In \citet{kim1996},
K-corrections were extended to handle additional cross-filter comparisons and
address zero-point corrections, but they started their derivation by
referencing the incorrect equation from \citet{oke1968}. The derivation in
\citet{kim1996} forms the basis for modern implementations of K-corrections.
This history is explored in more depth in Section \ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect compounds for greater
distances. Up until \citet{riess1998}, we didn't have distant enough
observations for this error to matter much for cosmology models, but in the
late 1990s, it was clear that our measurements for distance and redshift were
not linear. This observation led to a model that utilized a cosmological
constant and dark energy in order to explain the non-linear distance-redshift
graph. We will show in Section \ref{sec:tension} that correcting the flaw
with K-corrections leads to a linear graph that does not need to rely on a
cosmological constant.

Since \citet{planck2020}, it has been evident that something was missing with
our understanding of cosmology. This research used an alternative technique
based on measurements of the cosmic microwave background to compute the Hubble
constant, but as summarized in \citet{perivolaropoulos2022}, this measurement
was incompatible with the Hubble constant measured from SnIa. As we will
discuss in Section \ref{sec:tension}, fixing the error with K-corrections
produces a measurement of the Hubble constant that is compatible with models
based on the cosmic microwave background.

@@

\begin{equation}
 F'\left(\frac{\lambda}{1+z}\right) = \frac{\lambda}{(1+z)hc} F\left(\frac{\lambda}{1+z}\right).
\end{equation}

@@

\begin{equation}
\begin{aligned}
\label{eq:redshifted_density}
  R'(\lambda) &= F'\left(\frac{\lambda}{1+z}\right) \times \frac{1}{(1 + z)^2} \\
  R(\lambda) &= \frac{\lambda}{(1+z)hc} \times F\left(\frac{\lambda}{1+z}\right) \times \frac{1}{(1 + z)^2} .
\end{aligned}
\end{equation}

We deliberately do not combine all of the factors of $1+z$ together because
this form is more natural to implement in code.

@@

\begin{equation}
\begin{aligned}
  K'_{xy} &= \frac{\mathcal{F}_x P'_x}{\mathcal{F}_y P'_y} \\
         &= \frac{P'_x \int F'(\lambda) S_x(\lambda) d\lambda}
                 {P'_y \int R'(\lambda) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times
              \frac{\int F'(\lambda) S_x(\lambda) d\lambda}
                   {\int F'(\frac{\lambda}{1+z}) \times \frac{1}{(1 + z)^2} S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1+z)^2 \times
              \frac{\int \frac{\lambda}{hc} F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{(1+z)hc} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \\
         &= \frac{P'_x}{P'_y} \times (1 + z)^2 \times
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} .
\end{aligned}
\end{equation}

@@

\begin{equation}
\begin{aligned}
\label{eq:K2k}
  K_{xy} &= 2.5\text{log}(K'_{xy}) \\
         &= 2.5\text{log}\left(
            \frac{P'_x}{P'_y} \times (1 + z)^2 \times
            \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                 {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda}\right) \\
         &= 2.5 \left(
            \text{log} \left( \frac{P'_x}{P'_y} \right)
            + \text{log}( {(1 + z)^2})
            + \text{log}\left( \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda}
            \right) \right) \\
         &= 2.5 \text{log} \left( \frac{P'_x}{P'_y} \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \right) \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \right)
            - P_x + P_y .
\end{aligned}
\end{equation}

@@

\begin{equation}
\begin{aligned}
  K_{xy} &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \right)
            - P_x + P_y \\
         &= 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \right)
            + 2.5 \text{log} \left( \int \frac{\lambda}{hc} \mathcal{Z}(\lambda) S_x(\lambda) d\lambda \right)
            - 2.5 \text{log} \left( \int \frac{\lambda}{hc} \mathcal{Z}(\lambda) S_y(\lambda) d\lambda \right) \\
         &= 2.5 \text{log} \left(
              \frac{\int \lambda \mathcal{Z}(\lambda) S_x(\lambda) d\lambda}
                   {\int \lambda \mathcal{Z}(\lambda) S_y(\lambda) d\lambda}
             \right)
            + 5 \text{log} (1 + z)
            + 2.5 \text{log} \left(
              \frac{\int \lambda F(\lambda) S_x(\lambda) d\lambda}
                   {\int \frac{\lambda}{1+z} F\left(\frac{\lambda}{1+z}\right) S_y(\lambda) d\lambda} \right) .
\end{aligned}
\end{equation}

@@

\begin{itemize}
  \item The sign is different for the first term. This will have a subtle but
  frustrating effect for real measurements. Typically the zero-point values are
  similar, so this term is close to 0. However, the term isn't always zero. As
  we observe more distant supernovae, it is useful to switch our observational
  filter to a filter that is sensitive to longer wavelengths. Each observation
  filter will have its own small bias.

  This effect by itself can alter estimates of $H_0$ by several km/s / Mpc.
  While we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we aren't terribly confident in the numbers because we
  were unable to correct for the zero-point correction errors or verify that
  these errors don't exist.
